---
title: "lab_01"
author: "Derek Willis"
date: "7/21/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Turn off scientific notation
options(scipen=999)
```

## About this lab

To complete this lab, you need to:
* run existing code as directed (look for **Task**).
* modify existing code as directed (look for **Task**).
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).
When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

Remember to add any new verbs or concepts to your reference notebook!

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
```

## Load Data

We're going to load the slice of WinRed contribution data from West Virginia.

**Task** Run the codeblock below to load the data, which is contained in an RDS file.

```{r}
wv_winred_contribs <- read_rds("lab_01.rds")
```

## Examine Data

**Task** Run the codeblock below to use glimpse to get a sense of the column names, types and sample data.
**Q1** How many rows and columns does this dataframe have?
**A1** There are 13,382 rows and 24 columns.

```{r}
glimpse(wv_winred_contribs)
```

## Answer questions

**Q2** Which West Virginia city provided the largest number of contributions? How many?
**A2** 
Martinsburg had the largest amount of contributions with 781.

```{r}
wv_winred_contribs %>%
  group_by(city) %>%
  summarise(
    count_contribs = n()
  ) %>%
  arrange(desc(count_contribs))
```
**Q3** Which West Virginia city accounted for the fewest contributions? How many?
**A3** 
There were 40 cities with only 1 contribution. Some of these cities/communities included are Auburn, Belle, Colliers, Mason and Newell.
```{r}
wv_winred_contribs %>%
  group_by(city) %>%
  summarise(
    count_contribs = n()
  ) %>%
  arrange((count_contribs))
```
**Q4** What's your best educated guess to explain why the city identified in question 2 had more contributions compared to the cities identified in question 3? Hint: you won't find the answer in the data, you'll need to do some quick web research. You do not need to write any code for this question, so no codeblock is provided.  
**A4** 
From a few quick Google searches I found out that the city with the largest number of contributions, Martinsburg, is a city of around 17,500 and that it is also the eighth-largest municipality in the state. I looked up a few of the cities with one contribution and they were all cities with very small populations. Auburn is an unincorporated community with a population of 96. Belle has a population of around 1,100. Foster has a population of around 1,000. These communities are more rural and I imagine that making political contributions is not high on their list of things to do. There is also just not as many people to contribute.
**Q5** What recipient got the largest total amount of money? How many contributions did they receive and what was the average donation?
**A5** 
The SAVE AMERICA JOINT FUNDRAISING COMMITTEE received the largest amount of contributions with 2,448 for a total of $66,972.89. The average amount contributed was $27.36.
```{r}
wv_winred_contribs %>%
  group_by(memo_text) %>%
  summarise(
    count_contribs = n(),
    total_amount = sum(amount),
    mean_amount = mean(amount),
    median_amount = median(amount),
    min_amount = min(amount),
    max_amount = max(amount)
  ) %>%
  arrange(desc(total_amount))
```

**Q6** What is that top recipient and what explains its popularity? Hint: this will require web research, you won't be able to answer this purely from the data set.  You do not need to write any code for this question, so no codeblock is provided.  
**A6** 
The top recipient is the SAVE AMERICA JOINT FUNDRAISING COMMITTEE, which was formed by Donald Trump after he lost the 2020 election. It has become a gathering spot for Republican contributions. It's popularity most likely stems from it being created by Trump and supported/promoted by Republican officials. A quick Google search on it takes you to a page asking for donations touting those who donate as "patriots" who will go down in history for saving the U.S. from Biden and the "radical left."

**Q7**. Produce a summary table that has the following information for West Virginia as a whole (the whole state, not by city):
* count of contributions
* total amount of all contributions
* the average contribution amount
* the median contribution amount
* the smallest contribution amount
* the largest contribution amount

Store it as a variable object called `wv_summary` so that you can examine it with the data viewer. What newsworthy nuggets jump out at you that you could imagine dropping into a news story? Write a few summary sentences with overall findings.

**A7**. 
The first thing that stood out to me is that the minimum amount donated was one cent. It makes me wondering if that was a choice made as a statement or an option to add a donation on to your total price somewhere. It's probably not a super interesting story but that's definitely something I'm interested in knowing more about. Also looking at the numbers for the state as a whole makes we wonder how the number of contributions and the amounts donated compares to other states of similar sizes to West Virginia and with similar backgrounds. It could also be interesting to look into how WV compares to the states that border it like Kentucky or Maryland.
```{r}
wv_winred_contribs %>%
  group_by(state) %>%
  summarise(
    count_contribs = n(),
    total_amount = sum(amount),
    mean_amount = mean(amount),
    median_amount = median(amount),
    min_amount = min(amount),
    max_amount = max(amount)
  ) %>%
  arrange(desc(total_amount))
```
```{r}
wv_summary <- wv_winred_contribs %>%
  group_by(state) %>%
  summarise(
    count_contribs = n(),
    total_amount = sum(amount),
    mean_amount = mean(amount),
    median_amount = median(amount),
    min_amount = min(amount),
    max_amount = max(amount)
  ) %>%
  arrange(desc(total_amount))
```