---
title: "Reverese Engineering Round 2"
author: "Abbi Ross, Ryan Mercado, Sununu Bah, Dylan Manfre"
date: "11/19/2022"
output: html_notebook
---
#Loading libraries
```{r}
library(tidyverse)
library(janitor)
library(lubridate)
library(sf)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Loading in the 2022 data for both candidates
```{r}
elrich_22_data <- read_csv("data/Elrich_Contribs.csv")
```
This data set contains 2,981 rows and 13 columns. 

```{r}
blair_22_data <- read_csv("data/Blair_Contribs.csv")
```
This data set contains 556 rows and 13 columns. 


#Loading in the 2018 data for both candidates
```{r}
elrich_18_data <- read_csv("data/Elrich_Contribs_18.csv")
```
This data set contains 4,010 rows and 13 columns. 

```{r}
blair_18_data <- read_csv("data/Blair_Contribs_18.csv")
```
This data set contains 881 rows and 13 columns. 


#Loading in the 2022 county data
```{r}
dem_22_county_results <- read_csv("data/Dem_Results_22.csv")
```
This data set contains 32 rows and 11 columns. 

We also noticed that for some reason this dataset doesn't include all of the offices that were up for grabs, including county executive which is what we are looking at for Blair and Elrich. The 2018 file includes it though. That could be an issue for us, so we are looking for it on the website in other spots.

```{r}
dem_22_precinct_results<- read_csv("data/Dem_Precincts_22.csv")
```
This data set contains 18,298 rows and 13 columns. 

```{r}
correct_22_precinct_res<- read_csv("data/new_22_precincts.csv")
```


#Loading in the 2018 county data
```{r}
dem_18_county_results <- read_csv("data/Dem_Results_18.csv")
```
Rows: 216 Columns: 11

One thing I noticed about this data set is that it doesn't include mail in ballots unlike the 2022 data. In 2022 the mail in/provisional are lumped together. This could lead to us need to tweak our original question on that information.

```{r}
dem_18_precinct_results<- read_csv("data/Dem_Precincts_18.csv")
```
Rows: 27791 Columns: 11

#Cleaning the campaign finances data

```{r}
elrich_22_contribs <- elrich_22_data %>%
  clean_names() %>%
  mutate(contribution_date = mdy(contribution_date)) %>%
  select(-x13)
```

```{r}
elrich_18_contribs <- elrich_18_data %>%
  clean_names() %>%
  mutate(contribution_date = mdy(contribution_date)) %>%
  select(-x13)
```

```{r}
blair_22_contribs <- blair_22_data %>%
  clean_names() %>%
  mutate(contribution_date = mdy(contribution_date)) %>%
  select(-x13)
```

```{r}
blair_18_contribs <- blair_18_data %>%
  clean_names() %>%
  mutate(contribution_date = mdy(contribution_date)) %>%
  select(-x13)
```

#Cleaning the county election data

```{r}
dem_22_county_clean <- dem_22_county_results %>%
  clean_names()
```

```{r}
dem_18_county_clean <- dem_18_county_results %>%
  clean_names()
```

```{r}
dem_22_precincts_clean <- dem_22_precinct_results %>%
  clean_names()
```

```{r}
dem_18_precincts_clean <- dem_18_precinct_results %>%
  clean_names()
```

#Loading Expenses Data

```{r}
blair_22_expenses <- read_csv("data/blair_expenses_2022.csv")
```

```{r}
blair_22_exp_clean <- blair_22_expenses %>%
  clean_names()
```

```{r}
elrich_22_expenses <- read_csv("data/elrich_expenses_2022.csv")
```

```{r}
elrich_22_exp_clean <- elrich_22_expenses %>%
  clean_names()
```

```{r}
election_results_22 <- read_csv("data/22_precincts.csv")
```

```{r}
election_results_22 <- election_results_22 %>%
  clean_names()
```

# Week Two Update:
We  feel relatively comfortable with the data that we were able to find and download this week. Cleaning it was not that time consuming and it seems that we have all of the information we will need, other than the 2022 county results data and some iffy things surrounding mail/absentee ballots. Our first two questions are about self-funding and top contributors, which at this point do not seem like they should be an issue. Our third question is supposed to look at early voting and mail in votes. Right now it looks like the second half of the problem could be an issue. If so, we can tweak the question to focus solely on early votes and dive deeper into that.

Our other questions are about campaign finances going toward ads and precinct level data, both of which we should be good to go on.

The data sets we are analyzing are for two filing periods, 2022 and 2018.

We chose these years because Elrich and Blair ran for the same position in both. The 2022 and 2018 campaign finance data sets are limited to any contributions made during that filing year. 

Other than the one data set not including all of the information we need, we're not seeing any major reasons that won't be able to complete our work. We did see where we will have to do something similar to our first project, where the state is identified by acronym instead of it's full name, so we will have to change that.

Ryan did try to create a 2022 county result sheet, but we cannot get it to load.

# Week 3 Update

If all is correct, we successfully ran the code for our first two questions this week. These questions looked at Blair and Elrich's self-funding when it came to their campaigns and their top contributors. We were a little shocked at first when we saw that Blair's top five contributors were all family members or himself. We started answering our questions and what this means, but we do still need to look up all of the people who donated to Elrich and see if we can find out who they are and if they are a significant connection.

We also looked ahead to question 3. This question might require some tweaking. It's abotu mail in ballots and those votes are grouped a little differently in the 2018 county data, where just provisional votes are listed. In the 2022 data there are mail in and provisional. We believe that we will just have combine the mail in and provisional columns for the '22 data, which seems simple enough.

#Week 4 Update

This week we completed question 5, where found out that Blair spent nearly $3.1 mil in 2022. His top 3 expense categories were Media, Salarties and Direct Mail by Mail House (R). For this question we also wanted to look at his media expenses, where we found that $1,773 went to online advertising, which seemed low. His highest was television at $2,140,453. This makes sense given a lot of his voting demographic could be reached through television. It is interesting to see how much more he spent only to lose a second election so closely.

Elrich spent significantly less than Blair, with $528,393.29One of his top media exprense was consulting fees, at around $32k which we found interesting. His biggest media expense was television though, the same as Blair

We also worked on question 4, where we were able to find the .gdb file for Montgomery County boundaries which we loaded up and got a quick map for. We are stuck though for how to overlay the precincts onto that map. We are using pre-lab 9 as an example at the precinct map for that lab is already a gdb gile. We thought maybe Tigris was the best option, but that seems like a backward step since we're already looking at one county and have it on a map.

Once we figure this out, we can put in out data and start answering questions about which part of the county voted for whom and what that says about the candidate's voters they attracted/demographics/etc.

We also got part of the way through question 3, where we got the Elrich and Blair data seperated and bound together. For next week we need to chat with Derek about figuring out how to bring in the 2022 data, since the files for this year don't have them included for some reason. Once we have that, we fill confident about moving forward with the code.

##Important Findings

Some of the most important information we gathered this week is the difference in how they advertised. Blair went all out on advertising on most fronts, such as with television. He spent 4x as much on tv ads than Elrich. It shows an interesting dynamic between the millionaire Blair and the incumbent Elrich. 
We know Blair has the money and the connections, but it's interesting to see that despite all of this it wasn't enough two campaigns in a row. Once we figure our map in Question 4, we are interested in looking at how where Blair had more votes looks like demographics wise.

# Question 1 Attempt 1

How much money did David Blair and Marc Elrich fund themselves in 2022 vs. 2018? Blair is a businessman and millionaire, so it makes sense that he’s funding his own campaign. But how much exactly is he putting into it, and how does it compare to the past election and to how much Elrich’s campaign is raising?

# Answer 1:
Blair contributed $6,386,000 to his own campaign in 2022, while he contributed $7,630,576 in 2018.

Elrich contributed $3,275 to his campaign in 2022, making himself the second largest contributor to his campaign that year. In 2018 he contributed $150 to himself.

```{r}
blair_22_contribs_self <- blair_22_contribs %>%
filter(contributor_name == "BLAIR  DAVID  THOMAS") %>%
group_by(contributor_name) %>%
summarize(total_blair = sum(contribution_amount))
```

```{r}
blair_18_contribs_self <- blair_18_contribs %>%
filter(contributor_name == "BLAIR  DAVID  THOMAS") %>%
group_by(contributor_name) %>%
summarize(total_blair = sum(contribution_amount))
```

```{r}
elrich_22_contribs_self <- elrich_22_contribs %>%
filter(contributor_name == "Elrich  Marc") %>%
group_by(contributor_name) %>%
summarize(total_elrich = sum(contribution_amount))
```

```{r}
elrich_18_contribs_self <- elrich_18_contribs %>%
filter(contributor_name == "Elrich  Marc") %>%
group_by(contributor_name) %>%
summarize(total_elrich = sum(contribution_amount))
```

#Question 2, Attempt #1:
Who were the top 5 contributors to Blair and Elrich in 2022? How about in 2018? What are their connections to the candidates?

#Answer Atempt 1:

Blair's Top 5 for 2018
1.) David  Blair- $10,532,216
2.) Alice M Blair $6,000
3.) Andrew Blair- $6,000
4.) Jennifer Blair- $6,000
5.) Thomas L.Blair- $6,000

Blair's Top 5 for 2022
1.) David  Thomas Blair- $6,386,000
2.) Nelson Carbonell- $6,195
3.) Assal  Ethan- $6,000
4.) Alice Marjorie Blair- $6,000
5.) Thomas Blair- $6,000

All but one of Blair's top contributors from both years was not a family member, or himself. This was both surprising and not surprising. Who better to support a self funded bid for office than wealthy family members? If our Google search was correct, Nelson Carbonell is a GW University alumni and the president and CEO of consulting firm, Nelson Carbonell & Associates.

Elrich's Top 5 for 2018
1.) Montgomery County Government- $1,227,980
2.) Edward Redish- $1,000
3.) Tsega Haile- $300
4.) Anne Kaiser- $300
5.) Eden Durbin- $250

Elrich's Top 5 for 2022
1.) Montgomery County Government- $836,776
2.) March ELrich- $3,275
3.) Alfred Clark- $500
4.) Margaret Dennis-  $500
5.) Peter Kahn- $500

Elrich's top contributor in both campaigns was the Montgomery County Gov., which makes sense given that he was a part of a matching funds program. In 2022, he was the second top contributor, which was a little surprising. His second top contributor in 2018 is a Research Professor & Professor Emeritus of physics at the University of Maryland.

```{r}
top_elrich_18 <- elrich_18_contribs %>%
  group_by(contributor_name) %>%
  summarize(total = sum(contribution_amount)) %>%
  arrange(desc(total))%>%
  head(5)
```

```{r}
top_elrich_22 <- elrich_22_contribs %>%
  group_by(contributor_name) %>%
  summarize(total = sum(contribution_amount)) %>%
  arrange(desc(total))%>%
  head(5)
```

```{r}
top_blair_22 <- blair_22_contribs %>%
  mutate(contributor_name = case_when(
    contributor_name == "BLAIR  DAVID  THOMAS" ~ "Blair  David  Thomas",
    TRUE ~ contributor_name
  )) %>%
  group_by(contributor_name) %>%
  summarize(total = sum(contribution_amount)) %>%
  arrange(desc(total)) %>%
  head(5)
```

```{r}
top_blair_18 <- blair_18_contribs %>%
  mutate(contributor_name = case_when(
    contributor_name == "BLAIR  DAVID  THOMAS" ~ "Blair  David  Thomas",
    TRUE ~ contributor_name
  )) %>%
  group_by(contributor_name) %>%
  summarize(total = sum(contribution_amount)) %>%
  arrange(desc(total)) %>%
  head(5)
```

#Looking ahead at question 3:

#Q3) 
David Blair got more early/election day votes in 2022. Elrich got more Mail-in votes in 2022. How did that compare to 2018? People want to know how Marc Elrich won both times, first by 72 voters in 2018 and then 32 in 2022. Where did Marc do well in the three categories: Early voting, election day, and mail-in votes. Did this sway the result at all?

#Question 3 Answer Attempt 1

While Blair had more early and election votes in 2022, he did not in 2018. His election day total in 2018 was 25,739 and his early votes total was 8,671. Elrich beat him by 161 election votes and 329 in early votes.



```{r}
correct_22_precinct_res <- correct_22_precinct_res %>%
  clean_names()
```

```{r}
 blair_18_votes <- dem_18_county_clean %>%
filter(candidate_name == "David Blair")
elrich_18_votes <-  dem_18_county_clean %>%
  filter(candidate_name == "Marc Elrich")
voting_totals_18 <-  bind_rows(blair_18_votes, elrich_18_votes)
```

#Q3 Answer Attempt 2 with New Data

``` {r} 
voting_totals_22 <- correct_22_precinct_res %>%
filter(candidate_name == "David T. Blair" | candidate_name == "Marc Elrich") %>%
group_by(candidate_name) %>%
summarize(total_votes = sum(early_votes+election_night_votes+mail_in_ballot_1_votes+provisional_votes+mail_in_ballot_2_votes))
```



#Question 4
4. Which parts of the county voted for Elrich and which voted for Blair based on precinct-level voting? Are there differences in demographics of those areas? What about income? 



```{r}
precincts <- read_csv("data/election_precincts.csv")
glimpse(precincts)
```

```{r}
moco_boundaries <- st_read("data/moco.gdb")
glimpse(moco_boundaries)
```

```{r}
moco_precincts_correct <- st_read("data/MontMD_2022/MontMD_2022.shp") %>%
  st_zm(drop = TRUE)
glimpse(moco_precincts_correct)
```

```{r}
moco_precincts <- st_read("data/election_precincts/election_precincts.shp")
glimpse(moco_precincts)
```

#Maps for Q4

```{r}
correct_22_precinct_res <- correct_22_precinct_res %>%
  clean_names()
```

```{r}
moco_22_pre_results_filtered <- correct_22_precinct_res %>%
filter(candidate_name == "David T. Blair" | candidate_name == "Marc Elrich")
moco_22_pre_results_filtered
```

```{r}
blair_results_22_filter <- moco_22_pre_results_filtered %>% 
  filter(candidate_name == "David T. Blair")
blair_results_22_filter
```

```{r}
blair_results_22_filter_total <-blair_results_22_filter%>% 
  mutate(total_votes_blair = early_votes + election_night_votes + mail_in_ballot_1_votes + provisional_votes + mail_in_ballot_2_votes)
blair_results_22_filter_total
```

```{r}
blair_results_joined_2022<- inner_join(moco_precincts_correct, blair_results_22_filter_total, by=c("VOTESPRE"="election_district_precinct"))
```

###Blair Map####
```{r}
ggplot() +
  geom_sf(data=blair_results_joined_2022, aes(fill=total_votes_blair)) +
  scale_fill_viridis_b(option="magma")+
  theme_minimal()
```
####Elirich map####
```{r}
moco_22_pre_results_filtered_elrich <-correct_22_precinct_res %>%
filter(candidate_name == "Marc Elrich")
moco_22_pre_results_filtered_elrich
```

```{r}
elrich_22_filtered_total <-moco_22_pre_results_filtered_elrich%>% 
  mutate(total_votes_elrich = early_votes + election_night_votes + mail_in_ballot_1_votes + provisional_votes + mail_in_ballot_2_votes)
elrich_22_filtered_total
```

```{r}
elrich_results_joined_2022_map<- inner_join(moco_precincts_correct, elrich_22_filtered_total, by=c("VOTESPRE"="election_district_precinct"))
```

```{r}
ggplot() +
  geom_sf(data=elrich_results_joined_2022_map, aes(fill=total_votes_elrich)) +
  scale_fill_viridis_b(option="magma")+
  theme_minimal()
```
###percentage_change***
```{r}
correct_22_precinct_res <- correct_22_precinct_res %>%
  clean_names() 
```

```{r}
correct_22_executive <- correct_22_precinct_res %>%
  filter(office_name == "County Executive") %>% 
  mutate(candidate_total = early_votes + election_night_votes + mail_in_ballot_1_votes + mail_in_ballot_2_votes + provisional_votes)
```

```{r}
combining_candidates <- correct_22_executive %>%
  group_by(election_district_precinct) %>%
  summarise(total_precinct_votes = sum(early_votes + election_night_votes + mail_in_ballot_1_votes + mail_in_ballot_2_votes + provisional_votes))
```

```{r}
exec_combined_joined <- left_join(correct_22_executive, combining_candidates)
exec_combined_joined
```

```{r}
exec_joined_percentages <- exec_combined_joined %>%
  mutate(candidate_percents = candidate_total/total_precinct_votes*100)
exec_joined_percentages
```

```{r}
pivot_candidates <- exec_joined_percentages %>%
  select(election_district_precinct, candidate_name, candidate_percents, total_precinct_votes) %>% 
  pivot_wider(names_from = candidate_name, values_from = candidate_percents, values_fill = 0) %>%
  clean_names()
```

```{r}
pivot_candidates_joined<- inner_join(moco_precincts_correct, pivot_candidates, by=c("VOTESPRE"="election_district_precinct")) %>%
  mutate(blair_elrich_diff = david_t_blair - marc_elrich)
```

#Creating precinct percentage maps
-negatives are where elrich won, so the dark

```{r}
ggplot() +
  geom_sf(data=pivot_candidates_joined, aes(fill=blair_elrich_diff)) +
  scale_fill_viridis_b(option="magma")+
  theme_minimal()
```


```{r}

pivot_candidates_joined<- inner_join(moco_precincts_correct, pivot_candidates, by=c("VOTESPRE"="election_district_precinct")) %>%
  mutate(elrich_blair_diff = david_t_blair - marc_elrich) %>% 
  filter(elrich_blair_diff >= 10)

pivot_candidates_joined

ggplot() +
  geom_sf(data=moco_precincts_correct) +
  geom_sf(data=pivot_candidates_joined, aes(fill=elrich_blair_diff)) +
  scale_fill_viridis_b(option="magma")+
  theme_minimal()
```
```

#Question 5

```{r}
blair_22_expenses <- read_csv("data/blair_expenses_2022.csv")
```


```{r}
blair_22_exp_clean_grouped <- blair_22_exp_clean %>%
  group_by(expense_category) %>%
  summarize(total_category = sum(amount)) %>%
  arrange(desc(total_category))
```


# Blair spent nearly $3.1 mil in 2022. His top 3 expense categories were Media, Salarties and Direct Mail by Mail House (R).

#dive into media
```{r}
blair_expenses_22_medias <- blair_22_expenses %>%
filter(expense category == "Media") %>%
group_by(expense_purpose) %>%
summarize(total_category = sum(amount)) %>%
arrange(desc(total_category))

blair_expenses_2022_medias
```
#Blair spent $1773.00 on online advertising

```{r}
elrich_exp_clean_grouped <- elrich_22_exp_clean %>%
group_by(expense_category) %>%
summarize(total_category = sum(amount)) %>%
arrange(desc(total_category))
```
# Elrich spent $528393.29 on media which is significantly less than Blair

```{r}
elrich_22_exp_clean %>%
filter(expense_category == "Media") %>%
group_by(expense_purpose) %>%
summarize(total_category = sum(amount)) %>%
arrange(desc(total_category))
```
# Elrich spent 32K on consulting fees.
